---
title: "Delivery tactics: CRS Microdata"
format: html
editor: visual
---

```{r}
source(here::here('scripts', 'library.R'))
# load(here('data', 'crs_data_selected.RData'))
```

# Import CRS datasets

I extracted files for each year between 2002 and 2021 from the QWIDS platform by the OECD (i.e. 20 files in total).

```{r import}

folder_path <- here('data', 'crs')

file_list <- list.files(path = folder_path, full.names = TRUE)

read_clean_and_process <- function(file_path) {
  data <- fread(file_path) |> 
    clean_names()
  return(data)
}

data_list <- lapply(file_list, read_clean_and_process)

data_combined <- rbindlist(data_list)

save(data_combined, file = here('data', 'crs_data_combined.RData'))

```

```{r select_vars}

data_selected <- data_combined |> 
  select(1:26, 39, 44:54, 57:59, 62:67, 74:77, 78) |> 
  mutate_if(is.character, str_trim)

save(data_selected, file = here('data', 'crs_data_selected.RData'))

```

# Manipulate data set

## Donor type

```{r donor_type_name}

# list of Donor types from tge DAC CRS Codes
dac_member_list <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 18, 20, 21, 22, 40, 50, 59, 61, 68, 69, 75, 76, 84, 301, 302, 701, 742, 801, 820, 918)

multilateral_list <- c(104, 807, 811, 812, 901, 902, 903, 905, 906, 907, 909, 910, 913, 914, 915, 921, 923, 926, 928, 932, 940, 944, 948, 951, 952, 953, 954, 956, 958, 959, 960, 962, 963, 964, 966, 967, 971, 974, 976, 978, 979, 980, 981, 982, 983, 988, 990, 992, 997, 1011, 1012, 1013, 1014, 1015, 1016, 1017, 1018, 1019, 1020, 1023, 1024, 1025, 1037, 1038, 1039, 1041, 1044, 1045, 1046, 1047, 1048, 1049, 1050, 1052, 1053, 1054, 1055, 1311, 1312, 1313, 1401, 1404, 1406, 1409)

non_dac_member_list <- c(26, 30, 45, 55, 62, 70, 72, 77, 82, 83, 87, 130, 133, 358, 543, 546, 552, 561, 566, 576, 611, 613, 732, 764, 765)

private_donor_list <- c(1601, 1602, 1603, 1604, 1605, 1606, 1607, 1608, 1609, 1610, 1611, 1612, 1613, 1614, 1615, 1616, 1617, 1618, 1619, 1620, 1621, 1622, 1623, 1624, 1625, 1626, 1627, 1628, 1629, 1630, 1631, 1632, 1633, 1634, 1635, 1636, 1637, 1638, 1639, 1640, 1641, 1642, 1643, 1644, 1645, 1646, 1647, 1648)

# Note: manually add the following donor_code, because not included in the DAC CRS Codelist: 59, 1054, 1053, 1055, 1404, 1409, 1648.

# create new variable donor_type_name
data_donor_type <- data_selected |> 
  mutate(donor_type_name = fct(case_when( # Donor types
      donor_code %in% dac_member_list ~ 'DAC Members',
      donor_code %in% multilateral_list ~ 'Multilateral Donors',
      donor_code %in% non_dac_member_list ~ 'Non-DAC Members',
      donor_code %in% private_donor_list ~ 'Private Donors'
    )), .after = donor_code)

```

## Agency type

```{r agency_type}

agency_crs_code <- read_xlsx(here('data', 'dac_crs_codes.xlsx'),
          sheet = 'Agency') |> 
  clean_names() |> 
  mutate_if(is.character, str_squish)

# Note: use anti_join to check missing agency names in agency_crs_code

# add new agency_type vars
data_agency_type <- data_donor_type |> 
  left_join(agency_crs_code |> 
              select(donorcode, agency_code, agency_type, agency_type_name = type_name_e), 
                        by = c(donor_code = 'donorcode', 
                               agency_code = 'agency_code')) |> 
  mutate(agency_type_name = fct(agency_type_name)) |> 
  relocate(c(agency_type, agency_type_name), .after = agency_name)

# add agency code and type for non-State providers (i.e. Multilateral and Private Donors)
data_agency_type <- data_agency_type |> 
  mutate(
    agency_type = case_when(
      donor_type_name %in% c('Multilateral Donors', 'Private Donors') & agency_code == 1 ~ 1,
      donor_type_name %in% c('Multilateral Donors', 'Private Donors') & agency_code != 1 ~ 2,
      donor_code == 12 & agency_code %in% c(23:28, 30) ~ 2, # UK
      donor_code == 50 & agency_code == 59 ~ 2, # Spain
      donor_code == 62 & agency_code %in% (23:24) ~ 2, # Croatia
      donor_code == 546 & agency_code == 15 ~ 1, # Israel
      donor_code == 546 & agency_code != 15 ~ 2,
      .default = as.numeric(agency_type)
      ),
    agency_type_name = fct(case_when(
      donor_type_name %in% c('Multilateral Donors', 'Private Donors') & agency_code == 1 ~ 'Main Aid Agencies (in terms of budget)',
      donor_type_name %in% c('Multilateral Donors', 'Private Donors') & agency_code != 1 ~ 'Other Extending Aid Agencies',
      donor_code == 12 & agency_code %in% c(23:28, 30) ~ 'Other Extending Aid Agencies',
      donor_code == 50 & agency_code == 59 ~ 'Other Extending Aid Agencies',
      donor_code == 62 & agency_code %in% (23:24) ~ 'Other Extending Aid Agencies',
      donor_code == 546 & agency_code == 15 ~ 'Main Aid Agencies (in terms of budget)',
      donor_code == 546 & agency_code != 15 ~ 'Other Extending Aid Agencies',      
      .default = as.character(agency_type_name))
    ))

rm(data_donor_type)
```

## Nature of submission

See OECD's [Reporting Directives](https://one.oecd.org/document/DCD/DAC/STAT(2020)44/FINAL/En/pdf). (Nature of Submission – 379, p.86)

```{r new_activity_reported}

data_initial_report <- data_agency_type |>
  mutate(
    new_activity_reported = as_factor(
      if_else(initial_report == 1, "Yes", "No")
    ),
    .after = initial_report
  )

rm(data_agency_type)
```

## Region

```{r}

data_region <- data_initial_report |>
  mutate(
    africa = if_else(region_name %in% c(
      "North of Sahara", "South of Sahara",
      "Africa", "Africa, regional"
    ),
    "Yes", "No"
    ),
    .after = region_name
  )

rm(data_initial_report)
```

## Income group

```{r}

```

## Flow type

```{r}

```

## Bi/Multi

```{r}

```

## Channel of delivery

```{r}

```

```{r manipulate}

 |>  
  mutate(recipient_income_group = fct_relevel( # Recipient Income Group
      income_group_name,
      c("LDCs", "Other LICs", "LMICs", "UMICs", 
        "MADCTs", "Part I unallocated by income"),
      .after = income_group_name)) |> 
  mutate(bi_multi_group = fct(case_when( # Bi-/Multi Donor Group
      bi_multi %in% c(1, 3, 7, 8),  ~ 'Bilateral',
      bi_multi %in% c(2, 4) ~ 'Multilateral',
      bi_multi == 6 ~ 'Private Sector'
    )), .after = bi_multi) |> 
  mutate(bi_multi_name = fct(case_when(
      bi_multi == 1,  ~ 'Bilateral',
      bi_multi == 3,  ~ 'Bilateral (core contributions to NGOs and other private bodies / PPPs)',
      bi_multi == 7,  ~ 'Bilateral (triangular co-operation)',
      bi_multi == 2,  ~ 'Multilateral (inflows)',
      bi_multi == 4,  ~ 'Multilateral (outflows)',
      bi_multi == 6,  ~ 'Private Sector (outflows)'
  )), .after = bi_multi_group) |> 
  mutate(flow_name = fct_relevel(
    flow_name,
    c("ODA Grants", "Equity Investment", "ODA Loans")
  ))
  

# Save
save(oecd_crdf_rp_2000_21, oecd_climate_2013_21,
  file = here("data", "oecd", "oecd_crdf_rp_2000_21.RData")
)

```

Check for NAs

Rbind year data sets

Test merging with climate-related data set by oecd id, only selecting climate-relevant vars.
